Supplementary Material for Food-web complexity alters the fitness landscape of an insect herbivore

### 2018-06-08 supp_mat_complexity_selection.Rmd
substantial progress in visualizing univariate fitness landscapes. Multivariate is next!

## Calculate confidence intervals of relative fitness
relative_fitness_intervals <- function(logistic_model, newdata, bootstraps){
  model_matrix <- model.matrix(~ sc.gall_size + sc.clutch_size + sc.female_preference, newdata)

  get_absolute_fitness <- function(.) {
  absolute_fitness <- inverse_logit(model_matrix %*% fixef(.))
  absolute_fitness#/mean(absolute_fitness) 
  }

  get_bootstraps <-bootMer(logistic_model, FUN = get_absolute_fitness, nsim=bootstraps, parallel="multicore", ncpus=32)
  browser()

  get_intervals <- apply(get_bootstraps$t, 2, function(x) x[order(x)][c(round(bootstraps*0.025,0),round(bootstraps*0.975,0))])

  #mean_fitness <- mean(get_absolute_fitness(logistic_model))

  return(data.frame(relative_fitness = get_absolute_fitness(logistic_model),#/mean_fitness,
                    lower_2.5 = get_intervals[1, ],#/mean_fitness,
                    upper_97.5 = get_intervals[2, ],#/mean_fitness,
                    newdata))
}
```

@@

# Confidence intervals for selection gradients ----
boot_Beta_control <- bootMer(glmer_control, FUN = Beta_avg_grad, nsim=1000, parallel="multicore", ncpus=32)

@@

## Confidence intervals for selection gradients ----
boot_Beta_treatment <- bootMer(glmer_treatment, FUN = Beta_avg_grad, nsim=1000, parallel="multicore", ncpus=32)

@@

Illustrating the relationship between phenotype and mean fitness depends critically on the range of parameters shown. Perhaps this is what it is a good idea to show it based on the actual phenotypic distributions, or for 1 +/- SD

@@

RF_treatment_gall_size <- relative_fitness_intervals(
  logistic_model = glmer_treatment, 
  newdata = expand.grid(
     sc.gall_size = treatment_df$sc.gall_size, 
     sc.clutch_size = 0,
     sc.female_preference = 0),
  bootstraps=10)

@@

RF_control_gall_size <- relative_fitness_intervals(
  logistic_model = glmer_control, 
  newdata = expand.grid(
     sc.gall_size = treatment_df$sc.gall_size,
     sc.clutch_size = 0,
     sc.female_preference = 0),
  bootstraps=100)

>>>>

## Bootstrap fitness estimates
bootstrap_fitness <- function(logistic_model, newdata, bootstraps, intervals=c(0.025,0.975)){
  model_matrix <- model.matrix(~ sc.gall_size + sc.clutch_size + sc.female_preference, newdata)

  get_absolute_fitness <- function(.) inverse_logit(model_matrix %*% fixef(.))

  get_bootstraps <-bootMer(logistic_model, FUN = get_absolute_fitness, nsim=bootstraps, parallel="multicore", ncpus=32)

  get_intervals <- apply(get_bootstraps$t, 2, function(x) x[order(x)][c(round(bootstraps*intervals[1],0),round(bootstraps*intervals[2],0))])

  return(data.frame(newdata, 
                    average = get_absolute_fitness(logistic_model), 
                    lower = get_intervals[1, ],
                    upper = get_intervals[2, ],
                    t(get_bootstraps$t)))
}
```

@@

# Confidence intervals for selection gradients ----
boot_Beta_control <- bootMer(glmer_control, FUN = Beta_avg_grad, nsim=100, parallel="multicore", ncpus=32)

@@

## Confidence intervals for selection gradients ----
boot_Beta_treatment <- bootMer(glmer_treatment, FUN = Beta_avg_grad, nsim=100, parallel="multicore", ncpus=32)

@@

Illustrating the relationship between phenotype and mean fitness depends critically on the range of parameters shown. Perhaps this is what it is a good idea to show it based on the actual phenotypic distributions, or for 1 +/- SD.

The reason why the intervals look so chunky is because I've used the actual phenotypic distributions rather than sampling evenly at high densities across the distribution.

One interesting thing is that I don't believe the relative fitness intervals correspond to the absolute fitness intervals. This is likely because there can be a high absolute fitness, but the relative fitness might not change all that much...This means I need to figure out another way to plot these intervals. One easy way to do this would be to just return a list of two data frames, one that corresponds to absolute fitness and the other to relative fitness. This would be the easiest solution that would also remove a lot of redundancy in my code.

```{r Gall Size Fitness Landscape}

## Control ----
RF_control_gall_size <- bootstrap_fitness(
  logistic_model = glmer_control, 
  newdata = expand.grid(
     sc.gall_size = seq(-2,2,length.out=1000), #sort(control_df$sc.gall_size),
     sc.clutch_size = 0,
     sc.female_preference = 0),
  bootstraps=100)

## Treatment ----
RF_treatment_gall_size <- bootstrap_fitness(
  logistic_model = glmer_treatment, 
  newdata = expand.grid(
     sc.gall_size = seq(-2,2,length.out=1000), #sort(treatment_df$sc.gall_size), 
     sc.clutch_size = 0,
     sc.female_preference = 0),
  bootstraps=100)

## Spaghetti Plot: Relative Fitness ----   
bind_rows(mutate(RF_treatment_gall_size, Food_Web = "Simple"), 
          mutate(RF_control_gall_size, Food_Web = "Complex")) %>%
  select(-sc.clutch_size, -sc.female_preference) %>%
  gather(ID, fitness, -sc.gall_size, -Food_Web) %>%
  group_by(ID, Food_Web) %>%
  mutate(relative_fitness = fitness/mean(fitness)) %>%
  ungroup() %>%
  mutate(ID_group = ifelse(ID == "average", "average", "replicate")) %>% 
  ggplot(., aes(x=sc.gall_size, y=relative_fitness, group=interaction(ID,Food_Web), color=Food_Web, alpha=ID_group, size=ID_group)) +
  geom_line() +
  scale_alpha_manual(values=c(1,0.1)) +
  scale_size_manual(values=c(1.5,0.25))

## Interval Plot: Relative Fitness ----   
bind_rows(mutate(RF_treatment_gall_size, Food_Web = "Simple"), 
          mutate(RF_control_gall_size, Food_Web = "Complex")) %>%
  group_by(Food_Web) %>%
  mutate(relative_fitness = average/mean(average), lower_RF = lower/mean(lower), upper_RF = upper/mean(upper)) %>%
  ggplot(., aes(x=sc.gall_size, y=relative_fitness)) +
  geom_line(aes(color=Food_Web)) +
  geom_ribbon(aes(ymin=lower_RF, ymax=upper_RF, fill=Food_Web), alpha=0.25)

## Spaghetti Plot: Absolute Fitness ----   
bind_rows(mutate(RF_treatment_gall_size, Food_Web = "Simple"), 
          mutate(RF_control_gall_size, Food_Web = "Complex")) %>%
  select(-sc.clutch_size, -sc.female_preference) %>%
  gather(ID, fitness, -sc.gall_size, -Food_Web) %>%
  mutate(ID_group = ifelse(ID == "average", "average", "replicate")) %>% 
  ggplot(., aes(x=sc.gall_size, y=fitness, group=interaction(ID,Food_Web), color=Food_Web, alpha=ID_group, size=ID_group)) +
  geom_line() +
  scale_alpha_manual(values=c(1,0.1)) +
  scale_size_manual(values=c(1.5,0.25))

## Interval Plot: Absolute Fitness ----   
bind_rows(mutate(RF_treatment_gall_size, Food_Web = "Simple"), 
          mutate(RF_control_gall_size, Food_Web = "Complex")) %>%
  ggplot(., aes(x=sc.gall_size, y=average)) +
  geom_line(aes(color=Food_Web)) +
  geom_ribbon(aes(ymin=lower, ymax=upper, fill=Food_Web), alpha=0.25)
```

```{r Clutch Size Fitness Landscapes}

## Control ----
RF_control_clutch_size <- bootstrap_fitness(
  logistic_model = glmer_control, 
  newdata = expand.grid(
     sc.gall_size = 0,
     sc.clutch_size = sort(distinct(control_df, Plant_Position, Gall_Number, sc.clutch_size)$sc.clutch_size), # note that these aren't unique
     sc.female_preference = 0),
  bootstraps=100)

## Treatment ----
RF_treatment_clutch_size <- bootstrap_fitness(
  logistic_model = glmer_treatment, 
  newdata = expand.grid(
     sc.gall_size = 0,
     sc.clutch_size = sort(distinct(treatment_df, Plant_Position, Gall_Number, sc.clutch_size)$sc.clutch_size), # note that these aren't unique
     sc.female_preference = 0),
  bootstraps=100)

## Spaghetti Plot: Relative Fitness ----   
bind_rows(mutate(RF_treatment_clutch_size, Food_Web = "Simple"), 
          mutate(RF_control_clutch_size, Food_Web = "Complex")) %>%
  select(-sc.gall_size, -sc.female_preference) %>%
  gather(ID, fitness, -sc.clutch_size, -Food_Web) %>%
  group_by(ID, Food_Web) %>%
  mutate(relative_fitness = fitness/mean(fitness)) %>%
  ungroup() %>%
  mutate(ID_group = ifelse(ID == "average", "average", "replicate")) %>% 
  ggplot(., aes(x=sc.clutch_size, y=relative_fitness, group=interaction(ID,Food_Web), color=Food_Web, alpha=ID_group, size=ID_group)) +
  geom_line() +
  scale_alpha_manual(values=c(1,0.1)) +
  scale_size_manual(values=c(1.5,0.25))

## Interval Plot: Relative Fitness ----

# unclear to me why the complex intervals are weird...
bind_rows(mutate(RF_treatment_clutch_size, Food_Web = "Simple"), 
          mutate(RF_control_clutch_size, Food_Web = "Complex")) %>%
  group_by(Food_Web) %>%
  mutate(relative_fitness = average/mean(average), lower_RF = lower/mean(lower), upper_RF = upper/mean(upper)) %>%
  ggplot(., aes(x=sc.clutch_size, y=relative_fitness)) +
  geom_line(aes(color=Food_Web)) +
  geom_ribbon(aes(ymin=lower_RF, ymax=upper_RF, fill=Food_Web), alpha=0.25)

## Spaghetti Plot: Absolute Fitness ----   
bind_rows(mutate(RF_treatment_clutch_size, Food_Web = "Simple"), 
          mutate(RF_control_clutch_size, Food_Web = "Complex")) %>%
  select(-sc.gall_size, -sc.female_preference) %>%
  gather(ID, fitness, -sc.clutch_size, -Food_Web) %>%
  mutate(ID_group = ifelse(ID == "average", "average", "replicate")) %>% 
  ggplot(., aes(x=sc.clutch_size, y=fitness, group=interaction(ID,Food_Web), color=Food_Web, alpha=ID_group, size=ID_group)) +
  geom_line() +
  scale_alpha_manual(values=c(1,0.1)) +
  scale_size_manual(values=c(1.5,0.25))

## Interval Plot: Absolute Fitness ----   
bind_rows(mutate(RF_treatment_clutch_size, Food_Web = "Simple"), 
          mutate(RF_control_clutch_size, Food_Web = "Complex")) %>%
  ggplot(., aes(x=sc.clutch_size, y=average)) +
  geom_line(aes(color=Food_Web)) +
  geom_ribbon(aes(ymin=lower, ymax=upper, fill=Food_Web), alpha=0.25)
```

```{r Female Preference Fitness Landscapes}

## Control ----
RF_control_female_preference <- bootstrap_fitness(
  logistic_model = glmer_control, 
  newdata = expand.grid(
     sc.gall_size = 0,
     sc.clutch_size = 0, 
     sc.female_preference = sort(distinct(control_df, Plant_Position, sc.female_preference)$sc.female_preference)), 
  bootstraps=100)

## Treatment ----
RF_treatment_female_preference <- bootstrap_fitness(
  logistic_model = glmer_treatment, 
  newdata = expand.grid(
     sc.gall_size = 0,
     sc.clutch_size = 0,
     sc.female_preference = sort(distinct(treatment_df, Plant_Position, sc.female_preference)$sc.female_preference)), 
  bootstraps=100)

## Spaghetti Plot: Relative Fitness ----   
bind_rows(mutate(RF_treatment_female_preference, Food_Web = "Simple"), 
          mutate(RF_control_female_preference, Food_Web = "Complex")) %>%
  select(-sc.gall_size, -sc.clutch_size) %>%
  gather(ID, fitness, -sc.female_preference, -Food_Web) %>%
  group_by(ID, Food_Web) %>%
  mutate(relative_fitness = fitness/mean(fitness)) %>%
  ungroup() %>%
  mutate(ID_group = ifelse(ID == "average", "average", "replicate")) %>% 
  ggplot(., aes(x=sc.female_preference, y=relative_fitness, group=interaction(ID,Food_Web), color=Food_Web, alpha=ID_group, size=ID_group)) +
  geom_line() +
  scale_alpha_manual(values=c(1,0.1)) +
  scale_size_manual(values=c(1.5,0.25))

## Interval Plot: Relative Fitness ----

# unclear to me why the complex intervals are weird...
bind_rows(mutate(RF_treatment_female_preference, Food_Web = "Simple"), 
          mutate(RF_control_female_preference, Food_Web = "Complex")) %>%
  group_by(Food_Web) %>%
  mutate(relative_fitness = average/mean(average), lower_RF = lower/mean(lower), upper_RF = upper/mean(upper)) %>%
  ggplot(., aes(x=sc.female_preference, y=relative_fitness)) +
  geom_line(aes(color=Food_Web)) +
  geom_ribbon(aes(ymin=lower_RF, ymax=upper_RF, fill=Food_Web), alpha=0.25)

## Spaghetti Plot: Absolute Fitness ----   
bind_rows(mutate(RF_treatment_female_preference, Food_Web = "Simple"), 
          mutate(RF_control_female_preference, Food_Web = "Complex")) %>%
  select(-sc.gall_size, -sc.clutch_size) %>%
  gather(ID, fitness, -sc.female_preference, -Food_Web) %>%
  mutate(ID_group = ifelse(ID == "average", "average", "replicate")) %>% 
  ggplot(., aes(x=sc.female_preference, y=fitness, group=interaction(ID,Food_Web), color=Food_Web, alpha=ID_group, size=ID_group)) +
  geom_line() +
  scale_alpha_manual(values=c(1,0.1)) +
  scale_size_manual(values=c(1.5,0.25))

## Interval Plot: Absolute Fitness ----   
bind_rows(mutate(RF_treatment_female_preference, Food_Web = "Simple"), 
          mutate(RF_control_female_preference, Food_Web = "Complex")) %>%
  ggplot(., aes(x=sc.female_preference, y=average)) +
  geom_line(aes(color=Food_Web)) +
  geom_ribbon(aes(ymin=lower, ymax=upper, fill=Food_Web), alpha=0.25)
```

@@

RF_treatment_gall_size <- relative_fitness_intervals(
  logistic_model = glmer_treatment, 
  newdata = expand.grid(
     sc.gall_size = sort(treatment_df$sc.gall_size), 
     sc.clutch_size = 0,
     sc.female_preference = 0),
  bootstraps=10)

@@

RF_control_gall_size <- relative_fitness_intervals(
  logistic_model = glmer_control, 
  newdata = expand.grid(
     sc.gall_size = treatment_df$sc.gall_size,
     sc.clutch_size = 0,
     sc.female_preference = 0),
  bootstraps=10)
